---
title: "LPW_Summary"
author: "Jordan Watson"
date: "August 29, 2017"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.width=8,message=FALSE,warning=FALSE)

library(tidyverse)
library(readxl)
library(broom)
library(mgcv)
library(forcats)
library(viridis)
library(rgdal)
library(maps)
library(mapdata)
library(PBSmapping)
library(marmap)
library(grid)
library(gridExtra)
library(trend)
library(ggthemes)
```

The map is wonky here but the separate #png should work. We can discuss the bounds you'd like but I'll need a better shapefile for Baranof if we want to zoom any closer.

```{r LPW_Map,fig.cap="Map of Alaska showing the location of Little Port Walter Field Station (red dot)."}
data(nepacLLhigh)

xmin <- -137
xmax <- -133.5
ymin <-  56
ymax <-  58

insetxmin <- -170
insetxmax <- -130
insetymin <- 50
insetymax <- 71

world <- clipPolys(nepacLLhigh, xlim=c(xmin,xmax),ylim=c(ymin,ymax), keepExtra=FALSE)

mymap1 <- getNOAA.bathy(lon1 = xmin, lon2 = xmax, lat1 = ymin, lat2 = ymax, resolution = 2)

my.map <- autoplot(mymap1,geom=c("raster"),coast=FALSE) +   
  geom_polygon(data=world,aes(x=X,y=Y,group=factor(PID))) + 
  geom_point(data=data.frame(long=-134.645,lat=56.38),aes(long,lat),color="red",size=3) + 
  theme_bw() + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  theme(legend.position="none")

#my.map

#  Create our extent rectangle
my.extent<-data.frame(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax)

#  This is just a data.frame with four points.
#my.extent

#  Let's plot the inset
my.inset <- ggplot() + 
  geom_polygon(data=nepacLLhigh,aes(x=X,y=Y,group=PID),fill="grey30") + 
  coord_map(xlim=c(insetxmin,insetxmax),ylim=c(insetymin,insetymax)) + 
  geom_rect(data = my.extent, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            alpha=0, 
            colour="red", 
            size = 1, 
            linetype=1) + 
  theme_bw() + 
  theme(panel.grid = element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        plot.margin=unit(c(0,0,0,0),"mm"),
        panel.border = element_blank()) +
  labs(x=NULL,y=NULL)

#my.inset

##png(file="Figures/LPW_Map.png",w=1800,h=1800, res=300)
grid.newpage()
v1<-viewport(width = 1, height = 1, x = 0.5, y = 0.5) #plot area for the main map
v2<-viewport(width = 0.2, height = 0.2, x = 0.88, y = 0.88) #plot area for the inset map
print(my.map,vp=v1) 
print(my.inset,vp=v2)
##dev.off()
```

\pagebreak

```{r LPW_Map_Unuk,fig.cap="Map of Alaska showing the location of Little Port Walter Field Station (red dot)."}
data(nepacLLhigh)

xmin <- -137
xmax <- -130.9
ymin <-  55.5
ymax <-  58

insetxmin <- -170
insetxmax <- -130
insetymin <- 50
insetymax <- 71

world <- clipPolys(nepacLLhigh, xlim=c(xmin,xmax),ylim=c(ymin,ymax), keepExtra=FALSE)

mymap1 <- getNOAA.bathy(lon1 = xmin, lon2 = xmax, lat1 = ymin, lat2 = ymax, resolution = 2)

my.map <- autoplot(mymap1,geom=c("raster"),coast=FALSE) +   
  geom_polygon(data=world,aes(x=X,y=Y,group=factor(PID))) + 
  geom_point(data=data.frame(long=-134.645,lat=56.38),aes(long,lat),color="red",size=3) + 
  geom_point(data=data.frame(long=-131.06,lat=56.08),aes(long,lat),color="red",size=3) + 
  theme_bw() + 
  xlab("Longitude") + 
  ylab("Latitude") + 
  theme(legend.position="none") + 
  annotate("text",x=-131.45,y=56.11,label = "Unuk R.",color="white") + 
  annotate("text",x=-134.35,y=56.4,label = "LPW",color="white")

#my.map

#  Create our extent rectangle
my.extent<-data.frame(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax)

#  This is just a data.frame with four points.
#my.extent

#  Let's plot the inset
my.inset <- ggplot() + 
  geom_polygon(data=nepacLLhigh,aes(x=X,y=Y,group=PID),fill="grey30") + 
  coord_map(xlim=c(insetxmin,insetxmax),ylim=c(insetymin,insetymax)) + 
  geom_rect(data = my.extent, 
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            alpha=0, 
            colour="red", 
            size = 1, 
            linetype=1) + 
  theme_bw() + 
  theme(panel.grid = element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        plot.margin=unit(c(0,0,0,0),"mm"),
        panel.border = element_blank()) +
  labs(x=NULL,y=NULL)

#my.inset

png(file="Figures/LPW_Map_w_Unuk.png",w=1800,h=1800, res=300)
grid.newpage()
v1<-viewport(width = 1, height = 1, x = 0.5, y = 0.5) #plot area for the main map
v2<-viewport(width = 0.2, height = 0.2, x = 0.88, y = 0.75) #plot area for the inset map
print(my.map,vp=v1) 
print(my.inset,vp=v2)
dev.off()
```

\pagebreak

```{r readData}
temp <- read_csv("Data/All_Recoveries_All_LPW_Chinook.txt") %>% 
  inner_join(read_excel("Data/Indicators.xlsx",sheet="Sheet1") %>% 
               dplyr::select(brood_year,stock,tag_code)) %>% 
  mutate(age=run_year-brood_year,
         recov=ifelse(recovery_location_code=="1F1NE109 10","lpw","non-lpw")) %>% 
  dplyr::select(run_year,recovery_id,stock,brood_year,tag_code,sex,age,length,weight)

survdat <- read_csv("Data/RMIS_Survival.csv") %>% 
  mutate(tag_code=paste0("0",Tagcode)) %>% 
  inner_join(read_excel("Data/Indicators.xlsx",sheet="Sheet1") %>% 
               dplyr::select(brood_year,stock,tag_code)) %>% 
  filter(stock!="KETA R 101-30") %>% 
  rename(Survival=surv)

spawn <- temp %>% 
  left_join(survdat %>% 
              dplyr::select(brood_year,tag_code,Survival)) %>% 
  mutate(concat=paste0(age,sex)) %>% 
  filter(stock=="UNUK R" & !is.na(sex) & concat%in%c("5F","6F","4M","5M","6M") & brood_year>1976 & brood_year<2012) %>% 
  mutate(weight=as.numeric(weight))

spawn <- temp %>% 
  left_join(survdat %>% 
              dplyr::select(brood_year,tag_code,Survival)) %>% 
  mutate(concat=paste0(age,sex)) %>% 
  filter(!is.na(sex) & concat%in%c("5F","6F","4M","5M","6M") & brood_year>1976 & brood_year<2012) %>% 
  mutate(weight=as.numeric(weight))

spawn <- spawn %>% 
  mutate(wt_flag=ifelse(!is.na(weight) & weight>50,1,0)) %>% 
  filter(wt_flag!=1) %>% 
  dplyr::select(-wt_flag)
```

\pagebreak

```{r lengthScatter_portrait,fig.cap="Chinook length by sex and age."}

#  First for all data points (it's messy)
p <- spawn %>% 
  mutate(sex=fct_recode(sex,"Male"="M",
                        "Female"="F"),
         fage=fct_recode(as.factor(age),"2-Ocean"="4","3-Ocean"="5","4-Ocean"="6"),
         agesex=paste(sex,fage,sep="; "),
         agesex=fct_expand(factor(agesex),"Female; 2-Ocean"),
         agesex=fct_relevel(agesex,"Male; 2-Ocean","Male; 3-Ocean","Male; 4-Ocean","Female; 2-Ocean")) 

p1 <- p %>%  
  ggplot(aes(brood_year,length)) + 
  geom_point(size=0.5) + 
  facet_wrap(~agesex,drop = FALSE,scales="free_y",ncol=2,dir="v") + 
  stat_smooth(se=FALSE,color="black",linetype=1) + 
  stat_smooth(se=FALSE,method="lm",color="black",linetype=2) + 
  ylab("Length (mm)") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5)) + 
  scale_x_continuous(breaks=seq(1977,2011),
                     labels=c("1977","","","","1981","","","","1985","","","","1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","",""))

g <- ggplotGrob(p1)
# get the grobs that must be removed
rm_grobs <- g$layout$name %in% c("panel-2-2", "strip-t-2-1")
# remove grobs
g$grobs[rm_grobs] <- NULL
g$layout <- g$layout[!rm_grobs, ]
## move axis closer to panel
g$layout[g$layout$name == "axis-b-1-2", c("t", "b")] = c(9, 9)
#  If you don't use scales=free_y in the facet, you'll want this.
#g$layout[g$layout$name == "axis-l-2-1", c("l", "r")] = c(7, 7)
#grid.newpage()
#grid.draw(g)

#  Now create the cleaner (final?) figure with average length in each year
p1 <- p %>%  
  group_by(brood_year,agesex) %>% 
  summarise(length=mean(length,na.rm=TRUE)) %>% ungroup %>%
  ggplot(aes(brood_year,length)) + 
  geom_point() + 
  facet_wrap(~agesex,drop = FALSE,scales="free_y",ncol=2,dir="v") + 
  stat_smooth(se=FALSE,color="black",linetype=1) + 
  stat_smooth(se=FALSE,method="lm",color="black",linetype=2) + 
  ylab("Length (mm)") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5)) + 
  scale_x_continuous(breaks=seq(1977,2011),
                     labels=c("1977","","","","1981","","","","1985","","","","1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","",""))

g <- ggplotGrob(p1)
# get the grobs that must be removed
rm_grobs <- g$layout$name %in% c("panel-2-2", "strip-t-2-1")
# remove grobs
g$grobs[rm_grobs] <- NULL
g$layout <- g$layout[!rm_grobs, ]
## move axis closer to panel
g$layout[g$layout$name == "axis-b-1-2", c("t", "b")] = c(9, 9)
#  If you don't use scales=free_y in the facet, you'll want this.
#g$layout[g$layout$name == "axis-l-2-1", c("l", "r")] = c(7, 7)
#grid.newpage()
#grid.draw(g)


#  If we add Chickamin into the figure
#  Note that one point with a small mean is driving the apparent upward trend in size.
#  If we filter lengths to be greater than 510, this outlier effect is removed.
p1 <- p %>% 
  filter(length>510) %>% 
  group_by(brood_year,agesex,stock) %>% 
  summarise(length=mean(length,na.rm=TRUE)) %>% ungroup %>%
  rename(Stock=stock) %>% 
  ggplot(aes(brood_year,length,color=Stock)) + 
  geom_point() + 
  facet_wrap(~agesex,drop = FALSE,scales="free_y",ncol=2,dir="v") + 
  stat_smooth(aes(color=Stock),se=FALSE,linetype=1) + 
  stat_smooth(aes(color=Stock),se=FALSE,method="lm",linetype=2) + 
  ylab("Length (mm)") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5),
        legend.position=c(0.75,0.90)) + 
  scale_x_continuous(breaks=seq(1977,2011),
                     labels=c("1977","","","","1981","","","","1985","","","","1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","","")) + 
  scale_color_grey()


g <- ggplotGrob(p1)
# get the grobs that must be removed
rm_grobs <- g$layout$name %in% c("panel-2-2", "strip-t-2-1")
# remove grobs
g$grobs[rm_grobs] <- NULL
g$layout <- g$layout[!rm_grobs, ]
## move axis closer to panel
g$layout[g$layout$name == "axis-b-1-2", c("t", "b")] = c(9, 9)
#  If you don't use scales=free_y in the facet, you'll want this.
#g$layout[g$layout$name == "axis-l-2-1", c("l", "r")] = c(7, 7)
grid.newpage()
grid.draw(g)

png("Length_at_age.png",width=6,height=6,units="in",res=120);grid.draw(g);dev.off()
```



```{r, summarise_linear_models_length}
#  Create a table of linear model outputs
mdat <- p %>%
  filter(length>510) %>% 
  group_by(agesex,brood_year,stock) %>% 
  summarise(length=mean(length,na.rm=TRUE)) %>% data.frame

#  Make individual models for exploring residuals
#lm1 <- (lm(log(length)~brood_year,data=mdat %>%  filter(agesex=="Female; 3-Ocean")))
#lm2 <- (lm(log(length)~brood_year,data=mdat %>%  filter(agesex=="Female; 4-Ocean")))
#lm3 <- (lm(log(length)~brood_year,data=mdat %>%  filter(agesex=="Male; 2-Ocean")))
#lm4 <- (lm(log(length)~brood_year,data=mdat %>%  filter(agesex=="Male; 3-Ocean")))
#lm5 <- (lm(log(length)~brood_year,data=mdat %>%  filter(agesex=="Male; 4-Ocean")))


p.coef <- mdat %>% 
  group_by(agesex,stock) %>% 
  do(tidy(lm(length~brood_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  arrange(stock,agesex) %>% 
  data.frame

p.sum <- mdat %>% 
  group_by(agesex,stock) %>% 
  do(glance(lm(length~brood_year,data=.))) %>% 
  data.frame %>% 
  dplyr::select(agesex,stock,adj.r.squared) %>% 
    arrange(stock,agesex) %>% 
  data.frame
  
p.sum %>% 
  inner_join(p.coef) %>% 
  dplyr::select(agesex,stock,-term,estimate,se=std.error,-statistic,p.value,rsq=adj.r.squared) %>% 
  mutate_if(is.numeric,round,2) %>% 
  write_csv("Length_at_age_stock_linear_regressions.csv")
```


```{r, summarise_gam_models_length}
#  Create a table of GAM outputs
gam1 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Male; 2-Ocean" & stock=="UNUK R"))
gam2 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Male; 3-Ocean" & stock=="UNUK R"))
gam3 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Male; 4-Ocean" & stock=="UNUK R"))
gam4 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Female; 3-Ocean" & stock=="UNUK R"))
gam5 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Female; 4-Ocean" & stock=="UNUK R"))
gam6 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Male; 2-Ocean" & stock!="UNUK R"))
gam7 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Male; 3-Ocean" & stock!="UNUK R"))
gam8 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Male; 4-Ocean" & stock!="UNUK R"))
gam9 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Female; 3-Ocean" & stock!="UNUK R"))
gam10 <- gam(length~s(brood_year),data=mdat %>%  filter(agesex=="Female; 4-Ocean" & stock!="UNUK R"))

g1 <- summary(gam1)
g2 <- summary(gam2)
g3 <- summary(gam3)
g4 <- summary(gam4)
g5 <- summary(gam5)
g6 <- summary(gam6)
g7 <- summary(gam7)
g8 <- summary(gam8)
g9 <- summary(gam9)
g10 <- summary(gam10)

#gam.check(gam1)
#gam.check(gam2)
#gam.check(gam3)
#gam.check(gam4)
#gam.check(gam5)

mdat %>% 
  group_by(agesex) %>% 
  do(tidy(gam(length~s(brood_year),data=.))) %>% 
  inner_join(data.frame(agesex=unique(mdat$agesex),
                        dev.expl=c(g1$dev.expl,g2$dev.expl,g3$dev.expl,g4$dev.expl,g5$dev.expl,
                                   g6$dev.expl,g7$dev.expl,g8$dev.expl,g9$dev.expl,g10$dev.expl))) %>% 
  dplyr::select(-term,-ref.df,-statistic) %>% 
  mutate(dev.expl=100*dev.expl) %>% 
  mutate_if(is.numeric,round,2) %>% 
  write_csv("GAM_Model_Output.csv")
```


```{r,fig.cap="Compare length at age by sex (i.e., lpw fish) with non-lpw fish, for which sex is not known."}
temp <- read_csv("Data/All_Recoveries_All_LPW_Chinook.txt") %>% 
  inner_join(read_excel("Data/Indicators.xlsx",sheet="Sheet1") %>% 
               dplyr::select(brood_year,stock,tag_code)) %>% 
  mutate(age=run_year-brood_year,
         recov=ifelse(recovery_location_code=="1F1NE109 10","lpw","non-lpw")) %>% 
  dplyr::select(run_year,recovery_id,stock,brood_year,tag_code,sex,age,length,weight)


spawn2 <- temp %>% 
  left_join(survdat %>% 
              dplyr::select(brood_year,tag_code,Survival)) %>% 
  mutate(concat=paste0(age,sex)) %>% 
  filter(stock=="UNUK R" & brood_year>1976 & brood_year<2012) %>% 
  mutate(weight=as.numeric(weight))

spawn2 <- temp %>% 
  left_join(survdat %>% 
              dplyr::select(brood_year,tag_code,Survival)) %>% 
  mutate(concat=paste0(age,sex)) %>% 
  filter(brood_year>1976 & brood_year<2012) %>% 
  mutate(weight=as.numeric(weight))


spawn2 <- spawn2 %>% 
  mutate(wt_flag=ifelse(!is.na(weight) & weight>50,1,0),
         sexna=ifelse(is.na(sex),"N",sex)) %>% 
  filter(wt_flag!=1) %>% 
  dplyr::select(-wt_flag) %>% 
  filter(age>3 & age<7) %>%
  mutate(fage=fct_recode(as.factor(age),"2-Ocean"="4","3-Ocean"="5","4-Ocean"="6"),
         agesex=paste(sexna,fage,sep="; ")) %>% 
  filter(agesex!="F; 2-Ocean")

spawn2 %>% 
  filter(age>3 & age<7) %>% 
  group_by(age,sexna) %>% 
  do(tidy(lm(length~brood_year,data=.))) %>% 
  filter(term!="(Intercept)")
```


```{r Length_by_brood_unuk}
spawn2 %>% 
  filter(stock=="UNUK R") %>% 
  mutate(sexna=fct_recode(sexna,"Males"="M","Females"="F","No sex ID"="N")) %>%
  ggplot(aes(brood_year,length)) + 
  facet_wrap(~fage,drop = FALSE,scales="free_y",ncol=2,dir="v") + 
  stat_smooth(aes(linetype=sexna,color=sexna),se=FALSE) + 
  stat_smooth(aes(linetype=sexna,color=sexna),se=FALSE,method="lm") + 
  ylab("Length (mm)") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5),
        legend.title=element_blank(),
        legend.position=c(0.65,0.25)) + 
  scale_x_continuous(breaks=seq(1977,2011),
                     labels=c("1977","","","","1981","","","","1985","","","","1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","",""))
```


```{r Length_by_brood_chick}
spawn2 %>% 
  filter(stock!="UNUK R") %>% 
  mutate(sexna=fct_recode(sexna,"Males"="M","Females"="F","No sex ID"="N")) %>%
  ggplot(aes(brood_year,length)) + 
  facet_wrap(~fage,drop = FALSE,scales="free_y",ncol=2,dir="v") + 
  stat_smooth(aes(linetype=sexna,color=sexna),se=FALSE) + 
  stat_smooth(aes(linetype=sexna,color=sexna),se=FALSE,method="lm") + 
  ylab("Length (mm)") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5),
        legend.title=element_blank(),
        legend.position=c(0.65,0.25)) + 
  scale_x_continuous(breaks=seq(1977,2011),
                     labels=c("1977","","","","1981","","","","1985","","","","1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","",""))
```


```{r,numbers_of_fish_unuk,fig.cap="Compare length at age by sex (i.e., lpw fish) with non-lpw fish, for which sex is not known."}
tally(spawn2 %>% 
        filter(stock=="UNUK R") %>% 
        group_by(sexna,fage,brood_year)) %>% 
  ungroup %>% 
  mutate(sexna=fct_recode(sexna,"Males"="M","Females"="F","No sex ID"="N")) %>% 
  ggplot(aes(factor(brood_year),n,fill=factor(sexna))) + 
  geom_bar(stat="identity") + 
  facet_wrap(~fage,ncol=1)  + 
  scale_fill_viridis(discrete=TRUE) + 
  ylab("Number of fish") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5),
        legend.title=element_blank(),
        legend.position=c(0.5,0.9))
```


```{r,numbers_of_fish_chick,fig.cap="Compare length at age by sex (i.e., lpw fish) with non-lpw fish, for which sex is not known."}
tally(spawn2 %>% 
        filter(stock!="UNUK R") %>% 
        group_by(sexna,fage,brood_year)) %>% 
  ungroup %>% 
  mutate(sexna=fct_recode(sexna,"Males"="M","Females"="F","No sex ID"="N")) %>% 
  ggplot(aes(factor(brood_year),n,fill=factor(sexna))) + 
  geom_bar(stat="identity") + 
  facet_wrap(~fage,ncol=1)  + 
  scale_fill_viridis(discrete=TRUE) + 
  ylab("Number of fish") + 
  xlab("Brood Year") + 
  theme_bw()  + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5),
        legend.title=element_blank(),
        legend.position=c(0.5,0.9))
```


\pagebreak




```{r agestructure by brood year,fig.cap="Stock proportion by age class and brood year. Note that no year class existed for brood year 2000."}
tempdat <- spawn %>% 
  mutate(sex=fct_recode(sex,"Male"="M",
                        "Female"="F"),
         fage=fct_recode(as.factor(age),"2-Ocean"="4","3-Ocean"="5","4-Ocean"="6"),
         agesex=paste(sex,fage,sep="; "),
         agesex=fct_expand(factor(agesex),"Female; 2-Ocean"),
         agesex=fct_relevel(agesex,"Male; 2-Ocean","Male; 3-Ocean","Male; 4-Ocean","Female; 2-Ocean")) %>% 
group_by(sex,brood_year,stock) %>% 
  summarise(`2-Ocean`=length(age[age==4])/n(),
            `3-Ocean`=length(age[age==5])/n(),
            `4-Ocean`=length(age[age==6])/n()) %>% 
  gather(age,value,-brood_year,-sex,-stock) %>% 
  data.frame %>% 
  right_join(expand.grid(brood_year=unique(.$brood_year),
                         sex=c("Male","Female"),
                         age=c("2-Ocean","3-Ocean","4-Ocean"))) %>% 
  mutate(agesex=paste(sex,age,sep="; "),
         agesex=fct_expand(factor(agesex),"Female; 2-Ocean"),
         agesex=fct_relevel(agesex,"Male; 2-Ocean","Male; 3-Ocean","Male; 4-Ocean","Female; 2-Ocean"))


p <- tempdat %>% 
  ggplot(aes(brood_year,value)) + 
  geom_line() + 
  geom_point() + 
  facet_wrap(~agesex,ncol=2,dir="v",drop=FALSE) + 
  theme_bw() + 
  xlab("Brood Year") + 
  ylab("Proportion of return") + 
  scale_fill_viridis(discrete=TRUE,name="") + 
  scale_x_continuous(breaks=seq(1977,2011),
                     labels=c("1977","","","","1981","","","","1985","","","","1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","","")) + 
  theme(axis.text.x = element_text(angle = 90,size=8,vjust=0.5)) 

g <- ggplotGrob(p)
# get the grobs that must be removed
rm_grobs <- g$layout$name %in% c("panel-2-2", "strip-t-2-1")
# remove grobs
g$grobs[rm_grobs] <- NULL
g$layout <- g$layout[!rm_grobs, ]
## move axis closer to panel
graphics.off()
g$layout[g$layout$name == "axis-b-1-2", c("t", "b")] = c(9, 9)
#  If you don't use scales=free_y in the facet, you'll want this.
#g$layout[g$layout$name == "axis-l-2-1", c("l", "r")] = c(7, 7)
grid.newpage()
grid.draw(g)





x11();tempdat %>% 
  filter(brood_year<2013) %>% 
  ggplot(aes(brood_year,value,fill=age)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~sex) + 
  theme_bw() + 
  xlab("Brood Year") + 
  ylab("Proportion of return") + 
  scale_fill_viridis(discrete=TRUE,name="") + 
  scale_x_continuous(breaks=seq(1983,2011),
                     labels=c("","1984","","","","1988","","","","1992","","","","1996","","","","2000","","","","2004","","","","2008","","",""))+ 
  geom_hline(yintercept=0.05)
#dev.off()
```

\pagebreak

```{r, beta_regression}

data("GasolineYield", package = "betareg")
gy_logit <- betareg(yield ~ batch + temp, data = GasolineYield)
gy_loglog <- betareg(yield ~ batch, data = GasolineYield, link = "loglog")

summary(gy_logit)
summary(gy_loglog)


par(mfrow=c(2,2))
plot(gy_logit)
plot(gy_loglog)

mystock <- "UNUK R"
b1 <- betareg(value ~ brood_year, data = tempdat %>% filter(age=="2-Ocean" & sex=="Male" & value>0 & stock==mystock))
b2 <- betareg(value ~ brood_year, data = tempdat %>% filter(age=="3-Ocean" & sex=="Male" & value>0 & stock==mystock))
b3 <- betareg(value ~ brood_year, data = tempdat %>% filter(age=="4-Ocean" & sex=="Male" & value>0 & stock==mystock))
b4 <- betareg(value ~ brood_year, data = tempdat %>% filter(age=="3-Ocean" & sex=="Female" & value>0 & stock==mystock))
b5 <- betareg(value ~ brood_year, data = tempdat %>% filter(age=="4-Ocean" & sex=="Female" & value>0 & stock==mystock))

summary(b1)
summary(b2)
summary(b3)
summary(b4)
summary(b5)

par(mfrow=c(2,2));plot(b1)
par(mfrow=c(2,2));plot(b2)
par(mfrow=c(2,2));plot(b3)
par(mfrow=c(2,2));plot(b4)
par(mfrow=c(2,2));plot(b5)

b4 <- betareg(value ~ brood_year, data = tempdat %>% filter(age=="3-Ocean" & sex=="Female" & value>0 & stock==mystock & brood_year>1978))
summary(b4)
sapply(c("logit", "probit", "cloglog", "cauchit", "loglog"),function(x) logLik(update(b4, link = x)))



```




Look at the trends in key age classes over time. 
There appears to have been a relatively high degree of variability in female sizes in the earlier years of the dataset, whereas more recently, that has disappeared. There does not appear to be a trend in the mean over time per say. 

We fit some basic AR1 time series processes to the data as well, but there wasn't a significant relationship at a lag of 1 (or any other). We would not necessarily expect age-structure to be temporally autocorrelated but visually it seemed that this may have been the case so we checked. 

Visually, it appears that males have more variability in age-structure over time than females. 
```{r age structure brood year2, fig.cap="Stock proportion by age class and brood year. Note that no year class existed for brood year 2000."}
#png("Figures/AgeStructure_BY_Line.png",width=10,height=7.5,units="in",res=300)
x11();tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 4","Age 5","Age 6")) %>% 
  ggplot(aes(brood_year,value)) + 
  geom_line() + 
  facet_wrap(~sex+age) + 
  geom_point() + 
  ggtitle("Brood year") + 
  theme_bw() + 
  xlab("Brood Year") + 
  ylab("Proportion of return")

dat <- tempdat %>% 
  data.frame %>% 
  mutate(age=ifelse(sex=="Male" & as.character(age)=="Age 4","Age 5",as.character(age)),
         concat=paste0(sex,age)) %>% 
group_by(brood_year,age,sex) %>% 
  summarise(newvalue=sum(value,na.rm=TRUE)) %>% 


cor(dat$value[dat$sex=="Male" & dat$age=="Age 6" & !is.na(dat$value)],
    dat$value[dat$sex=="Female" & dat$age=="Age 6" & !is.na(dat$value)])

cor(dat$value[dat$sex=="Male" & dat$age=="Age 5" & !is.na(dat$value)],
    dat$value[dat$sex=="Female" & dat$age=="Age 5" & !is.na(dat$value)])

p1 <- ggplot() + 
  geom_line(data=dat %>% filter(sex=="Male" & age=="Age 6"),aes(brood_year,newvalue),linetype=2) + 
  geom_line(data=dat %>% filter(sex=="Female" & age=="Age 6"),aes(brood_year,newvalue),linetype=1) + theme_bw() + ggtitle("Age 6")
  
p2 <- ggplot() + 
  geom_line(data=dat %>% filter(sex=="Male" & age=="Age 5"),aes(brood_year,newvalue),linetype=2) + 
  geom_line(data=dat %>% filter(sex=="Female" & age=="Age 5"),aes(brood_year,newvalue),linetype=1) + theme_bw() + ggtitle("Age 5")

x11();grid.arrange(p1,p2)  

newt




+ 
  scale_x_continuous(breaks=seq(197,2011),
                     labels=c("","1984","","","","1988","","","","1992","","","","1996","","","","2000","","","","2004","","","","2008","","","")) + 
  theme(axis.text.x = element_text(vjust=0.5,angle = 90))

x11();tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & brood_year<2009 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6")) %>% 
  ggplot(aes(brood_year,value)) + 
  facet_wrap(~sex+age,scales="free") + 
  ggtitle("Brood year") + 
  geom_smooth(se=FALSE) + 
  geom_smooth(method="lm",se=FALSE) + 
  theme_bw() + 
  xlab("Brood Year") + 
  ylab("Proportion of return") + 
  scale_x_continuous(breaks=seq(1983,2011),
                     labels=c("","1984","","","","1988","","","","1992","","","","1996","","","","2000","","","","2004","","","","2008","","","")) + 
  theme(axis.text.x = element_text(vjust=0.5,angle = 90))

thresh1 <- tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & brood_year<2000 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6"))  

thresh2 <- tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & brood_year>=2000 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6"))
  
x11();tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & brood_year<2009 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6")) %>% 
  ggplot(aes(brood_year,value)) + 
  facet_wrap(~sex+age,scales="free") + 
  ggtitle("Brood year") + 
  geom_smooth(data=thresh1,method="lm",se=FALSE) + 
  geom_smooth(data=thresh2,method="lm",se=FALSE) + 
  theme_bw() + 
  xlab("Brood Year") + 
  ylab("Proportion of return") + 
  scale_x_continuous(breaks=seq(1983,2011),
                     labels=c("","1984","","","","1988","","","","1992","","","","1996","","","","2000","","","","2004","","","","2008","","","")) + 
  theme(axis.text.x = element_text(vjust=0.5,angle = 90))


tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & brood_year<2008 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6")) %>% 
  data.frame %>% 
  group_by(sex,age) %>% 
  do(glance(lm(value~brood_year,data=.)))

d1 <- tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & brood_year<2008 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6")) %>% 
  data.frame

d1 %>% 
  group_by(sex,age) %>% 
  do(augment(gam(value~s(brood_year),data=.)))

summary(gam(value~s(brood_year),data=d1 %>% filter(sex=="Female" & age=="Age 5")))

#dev.off()
```

\pagebreak

Correlations between each of the year classes for males were really that strong. Most striking was perhaps the utter lack of correlation between ages 5 and 6, suggesting that perhaps fish aren't trading off betweeen 5 and 6 as much as they are between 4 and 6 or 4 and 5 (though that wording sounds perhaps a bit anthropomorphized).

```{r,echo=TRUE}
cor(tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 3" & !is.na(tempdat$value)],
    tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 4" & !is.na(tempdat$value)])
cor(tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 3" & !is.na(tempdat$value)],
    tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 5" & !is.na(tempdat$value)])
cor(tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 3" & !is.na(tempdat$value)],
    tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 6" & !is.na(tempdat$value)])

cor(tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 4" & !is.na(tempdat$value)],
    tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 5" & !is.na(tempdat$value)])
cor(tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 4" & !is.na(tempdat$value)],
    tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 6" & !is.na(tempdat$value)])

cor(tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 5" & !is.na(tempdat$value)],
    tempdat$value[tempdat$sex=="Male" & tempdat$age=="Age 6" & !is.na(tempdat$value)])
```

\pagebreak

The smoothed scatterplots of age structure above really emphasize the lack of a real trend in age structure over time. The regressions below illustrate the danger that could be made by trying to force a trend. If we look at a regression of age-structure by brood year over time for the females, we see a linear regression that suggests an overall decrease in age over time (about a 1.5% decline in average age of fish per year). However, we could do a step regression before and after 2000, which appears to be an inflection point, and we would see approximately a 1% decrease per year prior to 2000 and a 1% increase per year after 2000. Prior to 2000, there appears to be a more variable age structure, with a variance in the proportion of age 5 females going from 0.2 prior to 2000 to 0.005 proceeding 2000). The size of returns or broods may be a relevant path of future consideration. 

For regressions by sex, we include only age classes that accounted for at least 5% of the brood in one year. This leaves 5 and 6s only for females and all but 7s for males.

In the regressions below, the "estimate" value is the linear model slope value in units of "proportion." In the paragraph above, I referred to regression results in terms of percents, so you could just multiply the "estimate" value by 100. I've omitted regression results for the males because they just weren't interesting. 

```{r age structure regression brood year,echo=TRUE}
#  For females regress only ages 5 and 6
#  All brood years (since 1983)
spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & brood_year<2011) %>% 
  do(tidy(lm(age~brood_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & brood_year<2011) %>% 
  do(tidy(mk.test(.$age)))


#  Pre-2000
spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & brood_year<2011) %>% 
  do(tidy(lm(age~brood_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & brood_year<2011) %>% 
  do(tidy(mk.test(.$age)))

# Post 2000
spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & brood_year>2000 & brood_year<2011) %>% 
  do(tidy(lm(age~brood_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & brood_year>2000 & brood_year<2011) %>% 
  do(tidy(mk.test(.$age)))


#  Variance of pre and post-2000 brood year proportions of age 5 females.
var((tempdat %>% filter(sex=="Female" & age=="Age 5" & brood_year<2000))$value)
var((tempdat %>% filter(sex=="Female" & age=="Age 5" & brood_year>2000 & brood_year<2011))$value)

```

\pagebreak

Note that 1989 does not have any age 7 fish because of the way we filtered the data, starting with brood year 1983. Seeemed better to include this year for the sake of the 3-6 year olds but note that the distribution is not technically complete here. 

```{r age structure by run year,fig.cap="Stock proportion by age class and run year. Note that no year class existed for brood year 2000."}
tempdat <- spawn2 %>% 
  filter(!is.na(sex) & run_year<2014 & run_year>1988) %>% 
  group_by(run_year,sex) %>% 
  summarise(`Age 3`=length(age[age==3])/n(),
            `Age 4`=length(age[age==4])/n(),
            `Age 5`=length(age[age==5])/n(),
            `Age 6`=length(age[age==6])/n(),
            `Age 7+`=length(age[age>=7])/n()) %>% 
  gather(age,value,-run_year,-sex)

tempdat <- tempdat %>% 
  right_join(expand.grid(run_year=c(unique(tempdat$run_year)),
                         sex=c("M","F"),
                         age=c("Age 3","Age 4","Age 5","Age 6","Age 7+"))) %>% 
  mutate(sex=fct_recode(sex,"Male"="M",
                        "Female"="F"),
         age=fct_relevel(age,"Age 7+","Age 6","Age 5","Age 4","Age 3"))

#png("Figures/AgeStructure_RunYear_Barplot.png",width=10,height=7.5,units="in",res=300)
tempdat %>% 
  filter(run_year<2014) %>% 
  ggplot(aes(run_year,value,fill=age)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~sex) + 
  theme_bw() + 
  xlab("Run Year") + 
  ylab("Proportion of return") + 
  scale_fill_viridis(discrete=TRUE,name="") + 
  scale_x_continuous(breaks=seq(1989,2017),
                     labels=c("1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","","","","2013","","","","2017"))
#dev.off()
```

\pagebreak

Similar to the distributions by brood year, we see variability in the age-structure by run year over time but not the most convincing trend. The loess smoothers illustrate this point by showing what looks like a relaxing of the overall mean increase per year during the middle time period. 

```{r age structure run year2,fig.cap="Stock proportion by age class and run year. Note that no year class existed for brood year 2000."}
#png("Figures/AgeStructure_RunYear_Scatter.png",width=10,height=7.5,units="in",res=300)
tempdat %>% 
  mutate(concat=paste0(sex,age)) %>% 
  filter(age!="Age 7+" & run_year<2017 & concat!="FemaleAge 3" & concat!="FemaleAge 4") %>% 
  mutate(age=fct_relevel(age,"Age 3","Age 4","Age 5","Age 6")) %>% 
  ggplot(aes(run_year,value)) + 
  geom_line() + 
  facet_wrap(~sex+age) + 
  geom_point() + 
  geom_smooth() + 
  theme_bw() + 
  xlab("Run Year") + 
  ylab("Proportion of return") + 
  scale_x_continuous(breaks=seq(1989,2017),
                     labels=c("1989","","","","1993","","","","1997","","","","2001","","","","2005","","","","2009","","","","2013","","","","2017")) + 
  theme(axis.text.x = element_text(vjust=0.5,angle = 90))
#dev.off()
```

\pagebreak

So what do the numbers say? Run year regressions show a 1.2% decrease in age per year for females 5-6. If we break the regression down on either side of brood year 2000 (which corresponds to run year 2006), we see a significant decline in age (1.2% per year) prior to run year 2007 and no significant trend since then for the same females. 

For males, we see an overall though very slight decrease in age (0.1% per year) across all years. Prior to 2007 run year, there is no significant change and since 2007, there is a 2% per year (significant) decline. 

```{r age structure regression run year}
#  For females regress only ages 5 and 6
#  All run years (1989-2013)
spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7)) %>% do(tidy(lm(age~run_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

#  Pre-2000
spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & run_year<2007) %>% 
  do(tidy(lm(age~run_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

# Post 2000
spawn2 %>% 
  filter(sex=="F" & !age%in%c(3,4,7) & run_year>2007) %>% 
  do(tidy(lm(age~run_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))


#  All run years (1989-2013)
spawn2 %>% 
  filter(sex=="M" & !age%in%c(3,4,7)) %>% do(tidy(lm(age~run_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

#  Pre-2000
spawn2 %>% 
  filter(sex=="M" & !age%in%c(7) & run_year<2007) %>% 
  do(tidy(lm(age~run_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))

# Post 2000
spawn2 %>% 
  filter(sex=="M" & !age%in%c(7) & run_year>2007) %>% 
  do(tidy(lm(age~run_year,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(estimate=round(estimate,4))
```
